---
title: "Basic Useful R"
output: ioslides_presentation
---
<!--include=FALSE-->
```{r setup, include=FALSE}
# Basic setup and library import
knitr::opts_chunk$set(echo = FALSE)
library(knitr)
library(dplyr)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

Code chunk can be cusomized by the following parameters

- `include = FALSE` prevents code and results from appearing in the finished file. R Markdown still runs the code in the chunk, and the results can be used by other chunks.
- `echo = FALSE` prevents code, but not the results from appearing in the finished file. This is a useful way to embed figures.
- `message = FALSE` prevents messages that are generated by code from appearing in the finished file.
- `warning = FALSE` prevents warnings that are generated by code from appearing in the finished.
- `fig.cap = "..."` adds a caption to graphical results.

## Various code language
### Python
display code and result
```{python, echo = T}
a = 'hello python'
print(a)
```
### Bash
display result only 
```{bash}
ls
```

## Slide with R Output
```{r cars, echo = TRUE}
summary(cars)
```

## Slide with table
```{r}
kable(cars[1:5, ], caption = "a table")
```

<!--This is a comment-->
## Slide with Plot
```{r pressure}
plot(pressure)
```

## dplyr package
There are 4 key operations:

- mutate() adds new variables that are functions of existing variables
- select() picks variables based on their names.
- filter() picks cases based on their values.
- summarise() reduces multiple values down to a single summary.
- arrange() changes the ordering of the rows.

## dplyr - filter() and selet()
```{r,echo = TRUE}
starwars %>% 
  filter(species == "Droid") %>%
  select(name, ends_with("color"))
```

## dplyr - mutate() and arrange()
```{r, echo = T}
starwars %>% 
  mutate(name, bmi = mass / ((height / 100)  ^ 2)) %>%
  select(name:mass, bmi) %>%
  arrange(desc(mass))
```

## dplyr - summarise() and group_by()
```{r, echo=T}
starwars %>%
  group_by(species) %>%
  summarise(
    n = n(),
    mass = mean(mass, na.rm = TRUE), 
    mass_median = median(mass, na.rm = TRUE)
  ) %>%
  filter(n > 1)
```
